### 2015-05-01 paper.tex
another commit

\documentclass{llncs}
%
\begin{document}

\title{Mass surveillance and the lost art of keeping a secret}

\author{Theo Tryfonas \and Tom Crick \and Michael Carter \and Panagiotis Andriotis}
\institute{Cryptography Group, University of Bristol, UK etc. etc.\\
\email { theo.tryfonas@bristol.ac.uk }}

\maketitle

\begin{abstract}
In this paper we 
\end{abstract}

\section{Introduction}
\label{sec:Introduction}
xxx

\end{document}